### 2024-02-12 globals/metadata.tex
update

\author{Aliaume Lopez}%
       {University of Warsaw}%
       {ad.lopez@uw.edu.pl}%
       {https://orcid.org/0000-0002-1825-0097}%
       {Gaëtan Douéneau-Tabot}

@@

\Copyright{Copyright Holder Name}

%TODO mandatory; please add comma-separated list of keywords
\keywords{Dummy keyword} 
\category{} %optional, e.g. invited paper
\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website

@@

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\author{Aliaume Lopez}%
       {University of Warsaw}%
       {ad.lopez@uw.edu.pl}%
       {https://orcid.org/0000-0002-1825-0097}%
       {}

@@

\Copyright{Aliaume Lopez}

%TODO mandatory; please add comma-separated list of keywords
\keywords{Rational series, weighted automata, polyregular function, commutative input} 
\category{} %optional, e.g. invited paper
\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website

\ccsdesc[500]{Theory of Computation~Models of Computation}
\ccsdesc[500]{Theory of computation~Automata extensions}

@@

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{ICALP 2023}
\EventAcronym{ICALP}
\EventYear{2023}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-02-12 parts/transducers.tex
update

\begin{lemma}
    Let $f \colon \Sigma^* \to \Rel$ be a \kl{polyblind function}.
    Then, there exists a computable function $g \colon \Nat^p \to \Rel$,
    that is \kl{commutative} and such that.
\end{lemma}

\subsection{}

@@

>>>>

\subsection{Deciding $\Nat$-polyregular functions}

The goal is to prove \cref{decide-n-poly-general:thm}.
To that end, we will use factorisation forests.
\begin{definition}
    Forests of depth $d$ by induction.
\end{definition}

\begin{remark}
    Place the nested higman embedding on the forests, it is a well-quasi-ordering.
\end{remark}

\begin{definition}
    Counting functions. $\vcount{q}$ where $q$ is a pattern.
\end{definition}

\begin{definition}
    Divisibility of patterns.
\end{definition}

\begin{definition}
    Forest iteration, where $q$ is a pattern.
\end{definition}

\begin{theorem}[Simon, restated in ]
    For every $f \in \ZPoly[k]$,
    there exists a computable function $h \colon \Sigma^* \to F_d(M)$,
    and a function $g \colon F_d(M) \to \Rel$ that is a counting function
    such that.
\end{theorem}

\textbf{TODO: domains should have a good shape, i.e.,
    intersections of sentences of the shape
    this pattern has at most "X" matches, and should be
    downwards closed}.

\textbf{TODO: problem, there is a problem, it is not because
    the number of matches is unbounded, that we can
    iterate the shape itself no?}

\begin{lemma}
    A function $g \colon F_d(M) \to \Rel$ is $\Nat$-polyregular
    over a domain $D$ if and only if for every 
    pattern iteration, the function obtained is
    $\Nat$-polyregular (commutative).
\end{lemma}
\begin{proof}
    By induction on the domain $D$, and on the divisibility ordering
    of patterns.
    Base case, for all pattern iteration, the value of $g$ is bounded.
    Then, $g$ has finitely many outputs, and they are all in $\Nat$,
    we have concluded.

    Let $q_i$ be a maximal pattern of $g$. If $q_i$ cannot be iterated in $D$,
    then the value of $q_i$ is bounded by some constant, and we can rewrite it
    as counting "full" patterns?

    Otherwise, $q_i$ can be iterated, and therefore we can
    pump to prove that its coefficient must be strictly positive.
    Then, using some derivation magic, we can recurse.
\end{proof}

\begin{theorem}
    \label{decide-n-poly-general:thm}
    Let $f \in \ZPoly[k]$, the following are equivalent:
    \begin{enumerate}
        \item $f \in \NPoly$;
        \item $f \in \NPoly[k]$;
        \item for all $h$ commutative polyregular function,
            $f \circ h$ in $\NPoly$
        \item for all pattern iteration,
            $f(pattern)$ is a \kl{$\Nat$-rational polynomial}.
    \end{enumerate}
    Furthermore, the properties are decidable.
\end{theorem}
\begin{proof}
    (1) implies (2) is okay, because of generic results.
    (2) implies (3) implies (4) is completely fine.
    Now, let us prove that (4) implies $1$.
\end{proof}

@@

\begin{lemma}
    Let $f \colon \Sigma^* \to \Rel$ be a \kl{polyblind function}.
    Then, there exists a computable function $g \colon \Nat^p \to \Rel$,
    that is \kl{commutative} and such that the following are equivalent:
    \begin{itemize}
        \item $g$ is $\Nat$-polyregular
        \item $f$ is $\Nat$-polyregular
    \end{itemize}
\end{lemma}
